---
title: "Getting Started with HSGconnect"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with HSGconnect}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

`HSGconnect` provides a convenient interface to access files stored on
[SwitchDrive](https://drive.switch.ch) directly from R. It connects via WebDAV,
so you can browse folders, read data files, and import BFS survey data — all
without manually downloading anything.

The typical workflow looks like this:

1. **One-time setup** — store your credentials securely.
2. **Connect** — authenticate against SwitchDrive.
3. **Browse** — list available files and folders.
4. **Read** — import data directly into R as a data frame.

## Prerequisites

Before you start, you should have an **app-specific password** for SwitchDrive.
It is not advised to use your regular university login. To create one:

1. Log in to [drive.switch.ch](https://drive.switch.ch).
2. Go to **Settings → Security**.
3. Under *App passwords*, create a new password and copy it.

You will also need the following R packages installed (they are listed as
dependencies and will be installed automatically):

- `httr`, `xml2` — for WebDAV communication
- `haven` — for reading `.sav` files (SPSS)
- `readr` — for reading `.csv` files
- `readxl` — for reading `.xlsx` files

## Step 1: Store Your Credentials

Run `sd_setup()` once to save your SwitchDrive username and app password to
your `.Renviron` file. This ensures they are available in every future R session
without hardcoding them in your scripts (security risk!!).

```{r}
library(HSGconnect)

sd_setup()
#> Enter your Switchdrive username (email): max.mustermann@unisg.ch
#> Enter your app password: ****
#> Credentials saved to .Renviron.
```

After setup, R will restart so the new environment variables take effect.
You only need to do this once.

## Step 2: Connect to SwitchDrive

At the beginning of each session, you can load the package and connect:

```{r}
library(HSGconnect)

sd_connect()
#> Successfully connected to Switchdrive as nick.glaettli@unisg.ch
```

`sd_connect()` reads your credentials from the environment variables set by
`sd_setup()` and validates the connection with a WebDAV request. The connection
is cached internally, so you only need to call it once per session.

**Note:** If you forget to call `sd_connect()`, don't worry — any function that
needs a connection will call it automatically. So you can also skip this step 
in each session without any problems.

## Step 3: Browse Files

Use `sd_list_files()` to see what's available on your SwitchDrive:

```{r}
# List the root directory ("Alle Dateien")
sd_list_files()
#> [1] "BFS Daten/"  "Projekte/"  "readme.txt"

# List a specific folder
sd_list_files("BFS Daten")
#> [1] "01 EHA Absolventen/"    "02 SSEE Studenten/"
#> [3] "03 Verträge/"           "04 Importinformationen/"

# Go deeper
sd_list_files("BFS Daten/01 EHA Absolventen/01 Arbeitsdaten")
#> [1] "eb_2023_hsg.sav"  "eb_2023_all.sav"  "zb_2023_hsg.sav"
```

Folders are indicated by a trailing `/`. Paths with spaces are handled
automatically.

**Note:** This functions serves for your convenience when wanting to import 
data from SwitchDrive. If you are only interested in importing BFS data,
you can rely on the `read_BFS_data()` function, as it automatically accesses
the appropriate folders. So there is no need to manually navegate SwitchDrive.

## Step 4: Read Data

### BFS Survey Data

For BFS survey data, use `read_BFS_data()`. It knows the folder structure and
file naming conventions, so you only need to specify the survey parameters:

```{r}
# EHA Erstbefragung 2023, HSG only (default)
eha <- read_BFS_data()
#> Reading eb_2023_hsg.sav...
#> Import successful: 1245 rows, 187 columns.

# EHA Zweitbefragung 2023, all universities
eha_all <- read_BFS_data(first_wave = FALSE, HSG_only = FALSE)
#> Reading zb_2023_all.sav...
#> Import successful: 38210 rows, 192 columns.

# SSEE 2024, HSG only
ssee <- read_BFS_data(survey = "SSEE", year = 2024)
#> Reading ssee_2024_hsg.sav...
#> Import successful: 2310 rows, 95 columns.

# EHA by cohort year (adds 1 to get the survey year)
eha_cohort <- read_BFS_data(year = 2022, cohort = TRUE)
#> Reading eb_2023_hsg.sav...
#> Import successful: 1245 rows, 187 columns.
```

**Parameters at a glance:**

| Parameter    | Default  | Description                                         |
|:-------------|:---------|:----------------------------------------------------|
| `survey`     | `"EHA"`  | Survey type: `"EHA"` or `"SSEE"`                    |
| `first_wave` | `TRUE`   | EHA only: `TRUE` for Erstbefragung, `FALSE` for Zweitbefragung |
| `year`       | `2023`   | Survey year (or cohort year if `cohort = TRUE`)      |
| `cohort`     | `FALSE`  | If `TRUE`, `year` is interpreted as cohort year      |
| `HSG_only`   | `TRUE`   | `TRUE` for HSG respondents only, `FALSE` for all     |

### Any File on SwitchDrive

For files outside the BFS structure, use `read_sd()`. It supports `.sav`,
`.csv`, and `.xlsx` files:

```{r}
# Read an SPSS file
df <- read_sd("Projekte/survey_results.sav")

# Read a CSV (comma-separated)
df <- read_sd("Daten/export.csv")

# Read a CSV with semicolon delimiter (common in Swiss/German data)
df <- read_sd("Daten/export.csv", delim = "semicol")

# Read an Excel file, specific sheet
df <- read_sd("Projekte/report.xlsx", sheet = "Rohdaten")
```

Additional arguments (`...`) are passed to the underlying import functions
(`haven::read_sav()`, `readr::read_csv()`, `readr::read_csv2()`,
`readxl::read_excel()`), so you can use parameters like `skip`, `col_types`,
or `na` as usual.

## Quick Reference

```{r}
# Full workflow in 4 lines
library(HSGconnect)
sd_connect()
sd_list_files("BFS Daten/01 EHA Absolventen/01 Arbeitsdaten")
eha <- read_BFS_data(survey = "EHA", year = 2023, HSG_only = TRUE)
```

## Troubleshooting

**Connection fails with status 401:**
Your credentials are likely incorrect. Check that you are using an app-specific
password (not your university login). Run `sd_setup()` again to update them.

**File not found (status 404):**
Use `sd_list_files()` to verify the exact path and filename. Watch out for
spaces and special characters in folder names.

**"No username provided" error:**
Your `.Renviron` file is not loaded. Make sure you restarted R after running
`sd_setup()`. You can verify with `Sys.getenv("SWITCH_USER")`.
